---
title: "DogBreed_Analysis"
author: "HunorKis"
date: "2022-12-23"
output: html_document
---
#Background for the project

##Core idea:
I would like to investigate the most important aspects by which we decide to buy a dog. It is a broadly discussed in dog related communites that Hungaricum dogbreeds are less and less chosen. To learn more about this topic you might want to check the [Ki kutyája vagyok én?](https://port.hu/adatlap/film/mozi/ki-kutyaja-vagyok-en-ki-kutyaja-vagyok-en-whose-dog-am-i/movie-243298) - titled movie, that was recently premiered.

I want to investigate the most important factors that define a top level dog and then I would like to create a model about the most important predictors.
Finally, I hope that I will see these predictor values to be lower in Hungarian breeds and that would serv as some sort of explanation.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read the dataset + add packages: To read the data I have used a simple readr method with the links for the dataset available on Github.
```{r}
library(tidyverse)
library(dplyr)
library(car)


breed_traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv')
trait_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/trait_description.csv')
breed_rank_all <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_rank.csv')


```

 Getting the data that I want to analyse: First of all, I needed to combine the two datasets for easier handling. All the changes will be find in this chunk.
 
```{r}

#Combine dataframes:
df_main <- cbind( breed_traits, breed_rank_all[2:9]) 



#remove spaces from colnames:
colnames(df_main) <- gsub(" ", "_", colnames(df_main))
colnames(df_main) <- gsub("/", "_", colnames(df_main))


#change colnames where number:

colnames(df_main)[18] <- "Rank_2013"
colnames(df_main)[19] <- "Rank_2014"
colnames(df_main)[20] <- "Rank_2015"
colnames(df_main)[21] <- "Rank_2016"
colnames(df_main)[22] <- "Rank_2017"
colnames(df_main)[23] <- "Rank_2018"
colnames(df_main)[24] <- "Rank_2019"
colnames(df_main)[25] <- "Rank_2020"

#Main table: since there are too much ranks to be handled, I added an extra column with popularity rankings, I have also filtered those that appear only in the 2020 ranking
bins <-c(0, 35, 70 , 105, 140, Inf)

df_main <- df_main %>% 
  filter(!is.na(Rank_2013)) %>% 
  mutate(Rank_cat = cut(Rank_2020, breaks = bins, labels = c("HighestRanking", "GoodRanking", "MediumRanking", "MediumLowRanking", "LowRanking"))) %>% 
  mutate(Coat_Type = as.factor(Coat_Type))


```
 
#EA: In this section I will explore the dataset

##Top 10 most and least popular breed

```{r}
head(df_main, 10)

tail(df_main, 10)



```

##Check whether 

```{r}

df_main %>% 
  ggplot()+
  aes(x = Rank_2020, y = Rank_2013)+
  geom_point()+
  geom_smooth(method='lm')+
  theme_light()+
  labs(title = "The changes in the position of ranked dogs between 2013-2020")
```

#From here on, I would like to analyze the characteristics of the most popular and least popular dogs from 2020

```{r}
df_main %>% 
  ggplot()+
  aes(x = Coat_Type, fill = Rank_cat) +
  geom_bar()+
  theme_light()+
  labs(title = "The distribution of Coat_type by Rankings in 2020",y = NULL, x = NULL)
```
##Now we will check the metrics that could predict how would the dog interact with people and dogs.

```{r}
df_main %>% 
  ggplot()+
  aes(y = Affectionate_With_Family)+
  geom_boxplot()+
  facet_wrap(~ Rank_cat)+
  theme_light()+
  labs(title = "Affectionate_With_Family")

df_main %>% 
  ggplot()+
  aes(y = Good_With_Young_Children)+
  geom_boxplot()+
  facet_wrap(~ Rank_cat) +
  theme_light()+
  labs(title = "Good_With_Young_Children")

df_main %>% 
  ggplot()+
  aes(y = Good_With_Other_Dogs)+
  geom_boxplot()+
  facet_wrap(~ Rank_cat) +
  theme_light()+
  labs(title = "Good_With_Other_Dogs")

```

##Now I want to further investigate the coat type and popularity

```{r}
df_main %>% 
  ggplot()+
  aes(y = Shedding_Level, fill = Rank_cat)+
  geom_boxplot() +
  theme_light()+
  labs(title = "Shedding Level by ranks")

df_main %>% 
  ggplot()+
  aes(y = Drooling_Level, fill = Rank_cat)+
  geom_boxplot()+
  theme_light()+
  labs(title = "Drooling Level by ranks")
```


##Now I will check how other role related information can impact the judgement

```{r}
df_main %>% 
  ggplot()+
  aes(y= Watchdog_Protective_Nature, fill = Rank_cat)+
  geom_boxplot() +
  theme_light()+
  labs(title = "Watchdog qualities by ranks")

df_main %>% 
  ggplot()+
  aes(y= Openness_To_Strangers, fill = Rank_cat)+
  geom_boxplot()  +
  theme_light()+
  labs(title = "Openness to Strangers by ranks")

#Presumable there is a correlation between the protectiveness and openness_To_Strangers

df_main %>% 
  ggplot()+
  aes(x= Openness_To_Strangers, y = Watchdog_Protective_Nature)+
  geom_count()+
  theme_light()+
  labs(title = "Compare the Openness To Strangers with Watchdog qualities")
  

cor.test(df_main$Openness_To_Strangers, df_main$Watchdog_Protective_Nature )
```
Unfortunately, I could not find a strong correlation between the Openness_To_Strangers and Watchdog_traits..

##Finally I would like to check the Trainability and Adaptability

```{r}
df_main %>% 
  ggplot()+
  aes(y= Trainability_Level, fill = Rank_cat)+
  geom_boxplot() +
  theme_light()+
  labs(title = "Trainability level by rank")

df_main %>% 
  ggplot()+
  aes(y= Adaptability_Level, fill = Rank_cat)+
  geom_boxplot() +
  theme_light()+
  labs(title = "Adaptability level by rank")

#Here, I also wanted to see the interaction:

df_main %>%
  ggplot() +
  aes(x = Trainability_Level, y = Adaptability_Level) +
  geom_count()+
  theme_light()+
  labs(title = "Adaptability level vs. Trainability levels")

cor.test(df_main$Adaptability_Level, df_main$Trainability_Level )
```

#Building Models

```{r}
Ranking_Model_Complex <- lm(Rank_2020~ Affectionate_With_Family + Good_With_Young_Children + Good_With_Other_Dogs + Watchdog_Protective_Nature + Openness_To_Strangers + Adaptability_Level + Trainability_Level, data = df_main)

summary(Ranking_Model_Complex)


Ranking_Model_FamilyTraits <- lm(Rank_2020~ Affectionate_With_Family + Good_With_Young_Children + Good_With_Other_Dogs, data = df_main)

summary(Ranking_Model_FamilyTraits)

Ranking_Model_role <- lm(Rank_2020~ Watchdog_Protective_Nature + Openness_To_Strangers + Adaptability_Level + Trainability_Level, data = df_main)

summary(Ranking_Model_role)


anova(Ranking_Model_Complex, Ranking_Model_FamilyTraits)

anova(Ranking_Model_Complex, Ranking_Model_role)

anova( Ranking_Model_role, Ranking_Model_FamilyTraits)



```

By comparing the three models, I have decided to go with the 3 factors that are connected to FamilyTraits

#Choosen model + Assumption check:
##Summary of the model
```{r}
summary(Ranking_Model_FamilyTraits)
```

##Check for outliers
```{r}
plot(Ranking_Model_FamilyTraits, which = 5)
```

##Checking normality
```{r}
residuals <- residuals(Ranking_Model_FamilyTraits) 
hist(residuals)
```

#&Checking linearity
```{r}
fitted_values <- fitted(Ranking_Model_FamilyTraits)
plot(fitted_values, residuals, xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0)
```

##Checking Multicollinearity
```{r}

vifs <- vif(Ranking_Model_FamilyTraits)

vifs

```

#Looking at the data of Hungarian dogs

```{r}
df_hungarian <- df_main %>% 
  filter(Breed %in% c("Komondorok", "Kuvaszok", "Pulik")) 
  


df_hungarian[, c("Breed","Affectionate_With_Family", "Good_With_Young_Children", "Good_With_Other_Dogs", "Rank_2020")]
```